---
title: "Home"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

Here is the [Results](SimulationExamples.html) with the new code.

## Install AlphaSimHlpR from LucianoRogerio github user


```{r, eval = FALSE}
library(devtools)
install_github("LucianoRogerio/AlphaSimHlpR", force = T, ref = "MarninPkg")
```

## Call all the packages required 

Change the working directory to the zip paste decompressed.

```{r}
suppressMessages(library(AlphaSimHlpR))
suppressMessages(library(tidyverse))
library(dplyr); library(furrr); library(here)

select <- dplyr::select
mutate <- dplyr::mutate
```

## Run the simulations parameters

```{r}
library(here)
### Small VDP scheme
schemeDF <- read.csv(here::here("data","baselineScheme - Test.csv"),
                     header = T, stringsAsFactors = F, sep = ";")
schemeDF

# Small alternative VDP scheme, number of entries were increased per stage
NschemeDF <- read.csv(here::here("data","baselineScheme - Test2.csv"),
                     header = T, stringsAsFactors = F, sep = ";")
NschemeDF

bsp1 <- specifyBSP(schemeDF = schemeDF,
                   nTrainPopCycles = 4, nYrsAsCandidates = 3, maxTrainingPopSize = 150,
                   nChr = 4, effPopSize = 200, quickHaplo = F,
                   segSites = 165, nQTL = 100, nSNP = 50, genVar = 1500,
                   gxeVar = NULL, gxyVar = 1500, gxlVar = 750, gxyxlVar = 300,
                   meanDD = 0.23, varDD = 0.05, relAA = 0.5,
                   stageToGenotype = "PYT",
                   nParents = 100, nCrosses = 100, nProgeny = 5, nClonesToNCRP = 3,
                   phenoF1toStage1 = F, errVarPreStage1 = 17500,
                   useCurrentPhenoTrain = F,
                   nCyclesToKeepRecords = 2,
                   selCritPipeAdv = selCritIID,
                   selCritPopImprov =  selCritIID)

bsp2 <- specifyBSP(schemeDF = NschemeDF,
                   nTrainPopCycles = 4, nYrsAsCandidates = 3, maxTrainingPopSize = 150,
                   nChr = 4, effPopSize = 200, quickHaplo = F,
                   segSites = 165, nQTL = 100, nSNP = 50, genVar = 1500,
                   gxeVar = NULL, gxyVar = 1500, gxlVar = 750, gxyxlVar = 300,
                   meanDD = 0.23, varDD = 0.05, relAA = 0.5,
                   stageToGenotype = "PYT",
                   nParents = 100, nCrosses = 100, nProgeny = 5, nClonesToNCRP = 3,
                   phenoF1toStage1 = F, errVarPreStage1 = 17500,
                   useCurrentPhenoTrain = F,
                   nCyclesToKeepRecords = 2,
                   selCritPipeAdv = selCritIID,
                   selCritPopImprov =  selCritIID)
```


## Include the Marnin functions to parallel the simulation

```{r}
source(here::here("code","runBurnInSchemes.R"))
source(here::here("code","runSchemesPostBurnIn.R"))
```

## Burn in Simulation script

It is not required to run this part of the script, since the results are already save in the file *BurnIn_IITA_PS.rds* at the *output* directory.

```{r, eval = FALSE}
start <- proc.time()[3]
burnIn_IITA_PS <- runBurnInSchemes(bsp = bsp1,
                                   nBurnInCycles=4,
                                   selCritPop="parentSelCritBLUP",
                                   selCritPipe="productSelCritBLUP",
                                   iniFunc="initializeScheme",
                                   productFunc="productPipeline",
                                   popImprovFunc="popImprovByParentSel",
                                   nReplications = 8, nSimCores = 4,
                                   nBLASthreads = 1)
end <- proc.time()[3]; print(paste0(floor((end - start)/60), "mins ",
                                    floor(((end - start)%%60)), "s elapsed - bsp1"))
#saveRDS(burnIn_IITA_PS,file = here::here("output","BurnIn_IITA_PS.rds"))
rm(start); rm(end)
```


## Post BurnIn Simulation with the Alternative VDP - Increasing the number of entries per stage

When we try to increase the number of entries per stage at the post burn in simulation, it is resulting at the following error:
```
Error: Problem with `mutate()` column `SimOutput`.
ℹ `SimOutput = future_map2(...)`.
x Trying to select invalid individuals
Run `rlang::last_error()` to see where the error occurred.
```

Here is the code to run the post burn in simulation

```{r, error = TRUE}
burnIn_IITA_PS <- readRDS(here::here("output", "BurnIn_PS.rds"))
start <- proc.time()[3]
postBurnIn_IITA_PS <- runSchemesPostBurnIn(simulations = burnIn_IITA_PS,
                                           newBSP = bsp2,
                                           nPostBurnInCycles = 12,
                                           selCritPop = "parentSelCritBLUP",
                                           selCritPipe = "productSelCritBLUP",
                                           productFunc = "productPipelinePostBurnIn",
                                           popImprovFunc = "popImprovByParentSel",
                                           nSimCores = 4,
                                           nBLASthreads = 1)
end <- proc.time()[3]; print(paste0(floor((end - start)/60), "mins ",
                                    floor(((end - start)%%60)), "s elapsed - bspP"))

```

Here is the *runSchemesPostBurnIn* function, that is resulting the Bug. Let's keep this function open in a table in RStudio, so I could use the line number as a reference to keep digging.

```{r, eval = FALSE}
View(runSchemesPostBurnIn)
```

## Digging the bug

I start by creating the following objects and run the function *runSchemesPostBurnIn* line by line.

```{r}
burnIn_IITA_PS <- readRDS(here::here("output", "BurnIn_PS.rds"))

simulations = burnIn_IITA_PS
newBSP = bsp2
nPostBurnInCycles = 12
selCritPop = "parentSelCritBLUP"
selCritPipe = "productSelCritBLUP"
productFunc = "productPipelinePostBurnIn"
popImprovFunc = "popImprovByParentSel"
nSimCores = 4
nBLASthreads = 1

require(furrr); plan(multisession, workers = nSimCores)
options(future.globals.maxSize=+Inf); options(future.rng.onMisuse="ignore")
```

at the line 37 and 38 of the *runSchemesPostBurnIn* function the function future_map2 is applyed to separate the scenarios by each repetition and run in parallel based on the number of cores informed in the function *plan* by the *workers* argument at line 34 of the *runSchemesPostBurnIn* function.

To keep digging I created the following objects:

```{r}

SimRep <- 1
burnInSim <- simulations$burnInSim[[SimRep]]

```

Then I proceed to the lines 41 to 66 of the *runSchemesPostBurnIn* function without any issue.

```{r}

      if(!is.null(nBLASthreads)) { RhpcBLASctl::blas_set_num_threads(nBLASthreads) }
      cat("******", SimRep, "\n")

      # This CONTINUES where previous sims left off
      ## no initialize step
      ## Keep burn-in stage sim params "SP"
      SP<-burnInSim$SP
      ## specify a potentially new bsp object
      ## (keep checks stored in burn-in stage's bsp)
      if(!is.null(newBSP)){
        bsp<-newBSP; bsp$checks<-burnInSim$bsp$checks
        bsp[["burnInBSP"]]<-burnInSim$bsp
        # years are indexed starting at year==0,
        ## so for 10 burn-in cycles, max value should be 9, store for later
        bsp[["maxYearBurnInStage"]]<-max(burnInSim$records$stageOutputs$year)
      } else { bsp<-burnInSim$bsp }
      ## 'historical' records from burn-in
      records<-burnInSim$records
      ## override burn-in specified product and population improvement funcs
      bsp[["productPipeline"]] <- get(productFunc)
      bsp[["populationImprovement"]] <- get(popImprovFunc)
      bsp[["selCritPipeAdv"]] <- get(selCritPipe)
      bsp[["selCritPopImprov"]] <- get(selCritPop)

      # Post burn-in cycles
      cat("\n"); cat("Post burn-in cycles"); cat("\n")
```

Then we arrive at the loop that is related to the bug (lines 67 to 71), more precisely at the *bsp$productPipeline* function, which is named as *productPipelinePostBurnIn*.

```{r, error = TRUE}
      for (cycle in 1:nPostBurnInCycles){
        cat(cycle, " ")
        records <- bsp$productPipeline(records, bsp, SP)
        records <- bsp$populationImprovement(records, bsp, SP)
      }
      
```

The *productPipelinePostBurnIn* could be viewed here.

```{r, eval = FALSE}
source(here::here("code", "newProductPipeline.R"))
source(here::here("code", "ProductPipeline.R"))

View(productPipelinePostBurnIn)
```

Then I proceed running line by line the function *productPipelinePostBurnIn*, starting with the line 28 until line 41.

```{r}
source(here::here("code", "ProductPipeline.R"))
# Calculate the selection criterion. selCritPipeAdv has to be given in bsp
  candidates <- records$F1@id
  selCrit <- bsp$selCritPipeAdv(records, candidates, bsp, SP)

  # Make summary for the incoming F1s
  year <- max(records$stageOutputs$year)+1 # Add a year relative to last year
  nF1 <- bsp$nCrosses * bsp$nProgeny
  nGenoRec <- nInd(records$F1)
  # Analyze the most-recent F1s
  newF1Idx <- nGenoRec - nF1 + 1:nF1
  id <- records$F1[newF1Idx]@id
  records$stageOutputs <- records$stageOutputs %>%
      dplyr::bind_rows(stageOutputs(id=id, f1=records$F1, selCrit=selCrit, stage=0, year=year, bsp=bsp))
  # Will be added to the phenotype records
  toAdd <- list()
```

Then to keep moving forward to the loop in line 42, I created the following object with the first Stage name and jump to line 43. I did the loop by my self.

##### PYT Stage

```{r, error = TRUE}
stage <- bsp$stageNames[1]
stage

 # Make a summary for this stage
    id <- last(records[[stage]])$id[1:bsp$nEntries[stage]]
    records$stageOutputs<-records$stageOutputs %>%
      dplyr::bind_rows(stageOutputs(id=id, f1=records$F1, selCrit=selCrit,
                                            stage=which(bsp$stageNames==stage),
                                            year=year, bsp=bsp))
```

Here I found the first bug, on line 45. Line 45 was wrote to remove the checks from the *id* vector. I did not realize that, because I tough that the issue the ***NAs*** when I was trying to run with the new number of entries per stage.

old code
```
id <- last(records[[stage]])$id[1:bsp$nEntries[stage]]
```
new code to remove the check from the *id* vector of each stage
```
id <- last(records[[stage]])$id %>% .[!. %in% bsp$checks@id]
```

Then I keep going the looping with the new code


```{r}
stage <- bsp$stageNames[1]
stage

 # Make a summary for this stage
    id <- last(records[[stage]])$id %>% .[!. %in% bsp$checks@id]
    records$stageOutputs<-records$stageOutputs %>%
      dplyr::bind_rows(stageOutputs(id=id, f1=records$F1, selCrit=selCrit,
                                            stage=which(bsp$stageNames==stage),
                                            year=year, bsp=bsp))

if(which(bsp$stageNames==stage) == 1){
      # Stage 1 different: no phenotypes but full Pop-class
      # Use phenotypes to select the F1 going into Stage 1?
      if(bsp$phenoF1toStage1){ # Use phenotypes to choose what goes to Stage 1
        phenoF1 <- setPheno(records$F1[newF1Idx], varE=bsp$errVarPreStage1, onlyPheno=T, simParam=SP)
        indToAdv <- records$F1@id[nGenoRec - nF1 + (phenoF1 %>% order(decreasing=T))[1:bsp$nEntries[stage]] %>% sort]
      } else {
        # Do the F1 have genotypic values that could be used?
        if(selCrit[newF1Idx] %>% is.na %>% all){ # Choose at random
          indToAdv <- records$F1@id[nGenoRec - nF1 + sort(sample(nF1, bsp$nEntries[stage]))]
        } else { # Use selCrit
          indToAdv <- records$F1@id[nGenoRec - nF1 + (selCrit[newF1Idx] %>%
                                                        order(decreasing=T))[1:bsp$nEntries[stage]] %>% sort]
        }
      }
    } else { # Beyond stage 1
      # Don't allow checks to be advanced: use 1:bsp$nEntries[stage-1]
      id <- last(records[[stage]])$id %>% .[!. %in% bsp$checks@id]
      selCritToAdv <- selCrit[id]
      indToAdv<-selCritToAdv %>% sort(.,decreasing = T) %>% .[1:bsp$nEntries[stage]] %>% names
    }

    entries <- records$F1[indToAdv]
    varE <- bsp$gxyVar + (bsp$gxlVar + bsp$gxyxlVar + bsp$errVars[stage] / bsp$nReps[stage]) / bsp$nLocs[stage]
    # reps=1 because varE is computed above
    entries <- setPheno(entries, varE=varE, reps=1, simParam=SP)
    phenoRec <- phenoRecFromPop(entries, bsp, stage)
    # If provided, add checks to the population
    if(!is.null(bsp$checks) & bsp$nChks[stage] > 0){
      varE <- bsp$gxyVar + (bsp$gxlVar + bsp$gxyxlVar + bsp$errVars[stage] / bsp$chkReps[stage]) / bsp$nLocs[stage]
      chkPheno <- setPheno(bsp$checks[1:bsp$nChks[stage]], varE=varE, reps=1, simParam=SP)
      chkRec <- phenoRecFromPop(chkPheno, bsp, stage, checks=T)
      phenoRec <- dplyr::bind_rows(phenoRec, chkRec)
    }
    toAdd <- c(toAdd, list(phenoRec))

```


##### AYT Stage

```{r}
stage <- bsp$stageNames[2]
stage

 # Make a summary for this stage
    id <- last(records[[stage]])$id %>% .[!. %in% bsp$checks@id]
    records$stageOutputs<-records$stageOutputs %>%
      dplyr::bind_rows(stageOutputs(id=id, f1=records$F1, selCrit=selCrit,
                                            stage=which(bsp$stageNames==stage),
                                            year=year, bsp=bsp))
    
if(which(bsp$stageNames==stage) == 1){
      # Stage 1 different: no phenotypes but full Pop-class
      # Use phenotypes to select the F1 going into Stage 1?
      if(bsp$phenoF1toStage1){ # Use phenotypes to choose what goes to Stage 1
        phenoF1 <- setPheno(records$F1[newF1Idx], varE=bsp$errVarPreStage1, onlyPheno=T, simParam=SP)
        indToAdv <- records$F1@id[nGenoRec - nF1 + (phenoF1 %>% order(decreasing=T))[1:bsp$nEntries[stage]] %>% sort]
      } else {
        # Do the F1 have genotypic values that could be used?
        if(selCrit[newF1Idx] %>% is.na %>% all){ # Choose at random
          indToAdv <- records$F1@id[nGenoRec - nF1 + sort(sample(nF1, bsp$nEntries[stage]))]
        } else { # Use selCrit
          indToAdv <- records$F1@id[nGenoRec - nF1 + (selCrit[newF1Idx] %>%
                                                        order(decreasing=T))[1:bsp$nEntries[stage]] %>% sort]
        }
      }
    } else { # Beyond stage 1
      # Don't allow checks to be advanced: use 1:bsp$nEntries[stage-1]
      id <- last(records[[stage]])$id %>% .[!. %in% bsp$checks@id]
      selCritToAdv <- selCrit[id]
      indToAdv<-selCritToAdv %>% sort(.,decreasing = T) %>% .[1:bsp$nEntries[stage]] %>% names
    }
```

Then I find another bug at the line 66 of the function *productPipelinePostBurnIn*. That was selecting clones for the next cycle but using the clones from the same stage.

Old code
```
      id <- last(records[[stage]]])$id %>% .[!. %in% bsp$checks@id]
```
New code to 
```
      id <- last(records[[*bsp$stageNames[which(bsp$stageNames==stage) -1]*]]])$id %>% .[!. %in% bsp$checks@id]
```
##### AYT Stage

```{r}
stage <- bsp$stageNames[2]
stage

 # Make a summary for this stage
    id <- last(records[[stage]])$id %>% .[!. %in% bsp$checks@id]
    records$stageOutputs<-records$stageOutputs %>%
      dplyr::bind_rows(stageOutputs(id=id, f1=records$F1, selCrit=selCrit,
                                            stage=which(bsp$stageNames==stage),
                                            year=year, bsp=bsp))
    
if(which(bsp$stageNames==stage) == 1){
      # Stage 1 different: no phenotypes but full Pop-class
      # Use phenotypes to select the F1 going into Stage 1?
      if(bsp$phenoF1toStage1){ # Use phenotypes to choose what goes to Stage 1
        phenoF1 <- setPheno(records$F1[newF1Idx], varE=bsp$errVarPreStage1, onlyPheno=T, simParam=SP)
        indToAdv <- records$F1@id[nGenoRec - nF1 + (phenoF1 %>% order(decreasing=T))[1:bsp$nEntries[stage]] %>% sort]
      } else {
        # Do the F1 have genotypic values that could be used?
        if(selCrit[newF1Idx] %>% is.na %>% all){ # Choose at random
          indToAdv <- records$F1@id[nGenoRec - nF1 + sort(sample(nF1, bsp$nEntries[stage]))]
        } else { # Use selCrit
          indToAdv <- records$F1@id[nGenoRec - nF1 + (selCrit[newF1Idx] %>%
                                                        order(decreasing=T))[1:bsp$nEntries[stage]] %>% sort]
        }
      }
    } else { # Beyond stage 1
      # Don't allow checks to be advanced: use 1:bsp$nEntries[stage-1]
      id <- last(records[[bsp$stageNames[which(bsp$stageNames==stage) -1]]])$id %>% .[!. %in% bsp$checks@id]
      selCritToAdv <- selCrit[id]
      indToAdv<-selCritToAdv %>% sort(.,decreasing = T) %>% .[1:bsp$nEntries[stage]] %>% names
    }

    entries <- records$F1[indToAdv]
    varE <- bsp$gxyVar + (bsp$gxlVar + bsp$gxyxlVar + bsp$errVars[stage] / bsp$nReps[stage]) / bsp$nLocs[stage]
    # reps=1 because varE is computed above
    entries <- setPheno(entries, varE=varE, reps=1, simParam=SP)
    phenoRec <- phenoRecFromPop(entries, bsp, stage)
    # If provided, add checks to the population
    if(!is.null(bsp$checks) & bsp$nChks[stage] > 0){
      varE <- bsp$gxyVar + (bsp$gxlVar + bsp$gxyxlVar + bsp$errVars[stage] / bsp$chkReps[stage]) / bsp$nLocs[stage]
      chkPheno <- setPheno(bsp$checks[1:bsp$nChks[stage]], varE=varE, reps=1, simParam=SP)
      chkRec <- phenoRecFromPop(chkPheno, bsp, stage, checks=T)
      phenoRec <- dplyr::bind_rows(phenoRec, chkRec)
    }
    toAdd <- c(toAdd, list(phenoRec))

```
##### UYT Stage

```{r}
stage <- bsp$stageNames[3]
stage

 # Make a summary for this stage
    id <- last(records[[stage]])$id %>% .[!. %in% bsp$checks@id]
    records$stageOutputs<-records$stageOutputs %>%
      dplyr::bind_rows(stageOutputs(id=id, f1=records$F1, selCrit=selCrit,
                                            stage=which(bsp$stageNames==stage),
                                            year=year, bsp=bsp))
    
if(which(bsp$stageNames==stage) == 1){
      # Stage 1 different: no phenotypes but full Pop-class
      # Use phenotypes to select the F1 going into Stage 1?
      if(bsp$phenoF1toStage1){ # Use phenotypes to choose what goes to Stage 1
        phenoF1 <- setPheno(records$F1[newF1Idx], varE=bsp$errVarPreStage1, onlyPheno=T, simParam=SP)
        indToAdv <- records$F1@id[nGenoRec - nF1 + (phenoF1 %>% order(decreasing=T))[1:bsp$nEntries[stage]] %>% sort]
      } else {
        # Do the F1 have genotypic values that could be used?
        if(selCrit[newF1Idx] %>% is.na %>% all){ # Choose at random
          indToAdv <- records$F1@id[nGenoRec - nF1 + sort(sample(nF1, bsp$nEntries[stage]))]
        } else { # Use selCrit
          indToAdv <- records$F1@id[nGenoRec - nF1 + (selCrit[newF1Idx] %>%
                                                        order(decreasing=T))[1:bsp$nEntries[stage]] %>% sort]
        }
      }
    } else { # Beyond stage 1
      # Don't allow checks to be advanced: use 1:bsp$nEntries[stage-1]
      id <- last(records[[bsp$stageNames[which(bsp$stageNames==stage) -1]]])$id %>% .[!. %in% bsp$checks@id]
      selCritToAdv <- selCrit[id]
      indToAdv<-selCritToAdv %>% sort(.,decreasing = T) %>% .[1:bsp$nEntries[stage]] %>% names
    }

    entries <- records$F1[indToAdv]
    varE <- bsp$gxyVar + (bsp$gxlVar + bsp$gxyxlVar + bsp$errVars[stage] / bsp$nReps[stage]) / bsp$nLocs[stage]
    # reps=1 because varE is computed above
    entries <- setPheno(entries, varE=varE, reps=1, simParam=SP)
    phenoRec <- phenoRecFromPop(entries, bsp, stage)
    # If provided, add checks to the population
    if(!is.null(bsp$checks) & bsp$nChks[stage] > 0){
      varE <- bsp$gxyVar + (bsp$gxlVar + bsp$gxyxlVar + bsp$errVars[stage] / bsp$chkReps[stage]) / bsp$nLocs[stage]
      chkPheno <- setPheno(bsp$checks[1:bsp$nChks[stage]], varE=varE, reps=1, simParam=SP)
      chkRec <- phenoRecFromPop(chkPheno, bsp, stage, checks=T)
      phenoRec <- dplyr::bind_rows(phenoRec, chkRec)
    }
    toAdd <- c(toAdd, list(phenoRec))

```

Finishing the *productPipelinePostBurnIn* function

```{r}
  for(stage in bsp$stageNames){
    records[[stage]] <- c(records[[stage]], toAdd[which(bsp$stageNames==stage)])
  }

  # Remove old records if needed
  if (length(records[[2]]) > bsp$nCyclesToKeepRecords) records <- removeOldestCyc(records, bsp)
```

Running the code of improvement population, which select the progenitor, and make the crosses

```{r}
records <- bsp$populationImprovement(records, bsp, SP)
```

Then I runned all over again for more three times as a loop, from **line 251** of this RmarkDown file.

Finishing the *runSchemePostBurnIn* function for one replicate.

```{r}
      # Finalize the stageOutputs
      records <- AlphaSimHlpR:::lastCycStgOut(records, bsp, SP)

      postBurnIn_IITA_PS <- list(list(records=records,
                                      bsp=bsp,
                                      SP=SP))
```

## Examples running the new code
Here is new [examples](SimulationExamples.html)
